\chapter{Field Trip Questions}
\label{apx:ftquestions}
{\footnotesize
\textbf{Interview questions:}
\begin{enumerate}
    \item How many do you live with, and what age are they?
    \item Which energy resources do you use (like electricity, gas, oil or others)?
    \item What do you and household use electricity for?
    \item On the SHS, how long do you keep the lights on daily, what power setting, and when do you use them?
    \item On the SHS, how many hours do you charge your phone daily, and when do you do it?
    \item Does the battery ever run out, and what does it take to happen?
    \item Do you use the system for anything else than lighting and phone charging?
    \item What is your electricity price?
    \item What is your income situation?
    \item How did you choose where to put the system, and where is it?
    \item Has there been any issues with the system, and is it time consuming to maintain - if so how many hours?
    \item Is the system simple enough to use, or would you want more training?
    \item How has it affected life in the household?
    \item What are the best use cases you have experienced?
    \item What do you think a fair price for this product would be if you were to buy it now?
    \item Who would you recommend it to, if anyone?
    \item What are your thoughts on solar power as an energy source?
\end{enumerate}
\textbf{Added mid field trip:}
\begin{enumerate}
    \item How frequent and how long are power outages?
    \item Have you replaced candles after receiving the system?
\end{enumerate}
\textbf{Specifically noted observations:}
\begin{enumerate}
    \item System placements, tilt angle and azimuth
    \item Type of system
\end{enumerate}}